---
title: "Plot"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Diwali Sales Final Plot

```{r}
library(tidyverse)
library(tidytuesdayR)
library(hrbrthemes)
library(scales)
library(lubridate)
library(ggrepel)
library(patchwork)
library(here)
library(gganimate)
library(patchwork)
library(camcorder)
library(ragg)
library(clipr)
library(emo)
library(forcats)
library(tidytext)
library(tidylo)
library(gt)
library(gtExtras)

gg_record(
  dir = file.path(getwd(), "final-plot-process"), 
  device = "png", 
  units = "in",  
  dpi = 300    
)

# Define default path for saving plots
plot_path <- here::here("output", paste0("tidytuesday_", 2023, "_", 46))
if (!dir.exists(plot_path)) {
  dir.create(plot_path, recursive = TRUE)
}

```

Unfortunately this week, not all of the output is made in R. I would have liked to, but couldn't find the time to solve some of the layout issues I was having. As a result, the table is made with `gT` but the finished layout is done in Figma.

## Prepping the data

```{r}
# Extract the category part from the product_category string
extract_category <- function(category) {
  str_split(category, pattern = "___", simplify = TRUE)[, 1]
}

# Mapping function for categories to emojis
category_to_emoji <- function(category) {
  clean_category <- extract_category(category)
  switch(
    clean_category,
    "Beauty" = emo::ji("lipstick"),
    "Clothing & Apparel" = emo::ji("shirt"),
    "Electronics & Gadgets" = emo::ji("iphone"),
    "Food" = emo::ji("pizza"),
    "Footwear & Shoes" = emo::ji("athletic_shoe"),
    "Games & Toys" = emo::ji("video_game"),
    "Household items" = emo::ji("bulb"),
    "Sports Products" = emo::ji("basketball"),
    ""
  ) 
}

state_abbr <- function(state) {
  switch(
    as.character(state), 
    "Andhra Pradesh" = "AP",
    "Maharashtra" = "MH",
    "Delhi" = "DL",
    "Karnataka" = "KA",
    "Madhya Pradesh" = "MP",
    "Uttar Pradesh" = "UP",
    NA  
  )
}

data <- tidytuesdayR::tt_load(2023, week = 46)$diwali_sales_data %>%
  janitor::clean_names() %>%
  select(
    gender,
    age_group,
    age,
    marital_status,
    state,
    zone,
    occupation,
    product_category,
    amount
  ) %>%
  drop_na() %>%
  filter(fct_lump(state, 6) != "Other",
         fct_lump(product_category, 7) != "Other") %>%
  count(state, product_category) %>%
  complete(state, product_category, fill = list(n = 0)) %>%
  bind_log_odds(state, product_category, n) %>%
  mutate(state = fct_reorder(state,-n, sum)) %>%
  mutate(product_category = reorder_within(product_category, log_odds_weighted, state)) %>%
  mutate(
    emoji = sapply(product_category, category_to_emoji),
    category = paste(emoji, sapply(product_category, extract_category)),
    state_code = sapply(state, state_abbr),
    state_name = paste(state, '<span style="display: block; font-style:italic; color:#c3c3c3;">', as.character(state_code), '</span>', sep = "")
  ) %>%
  group_by(state) %>%
  mutate(perc = n / sum(n) * 100,
         inner_square_size = sqrt(perc / 100))

x_labels <- paste0(2 ^ abs(seq(-6, 9, 3)), "x")
x_labels <- ifelse(x_labels == "1x", "Same", x_labels)

```

## Creating the Table

```{r}

 data %>%
  ungroup() %>%
    arrange(state, desc(log_odds_weighted)) %>%
  select(
    state,
    category,
    n,
    perc,
    log_odds_weighted,
    inner_square_size,
    product_category,
    emoji
  ) %>%
  gt(groupname_col = 'state') %>%
  row_group_order(
    groups = c(
      "Andhra Pradesh",
      "Madhya Pradesh",
      "Delhi",
      "Karnataka",
      "Maharashtra",
      "Uttar Pradesh"
    )
  ) %>%
  tab_header(
    title = html(
      '<span style="width: 50%;">What sells best in Diwali across different states in India?</span>'
    ),
    subtitle = html(
      '<p style="width: 80%;">Footwear in Delhi, electronics in Andhra Pradesh; some types of products outrank others in their odds of topping sales during Diwali shopping season</p>'
    )
  ) %>%
  tab_source_note(
    source_note = gt::html(
      "Data shared on Kaggle by Saad Haroon, tidytuesday project <br> <span style='font-size: 10px;'>Graphic: Aman Bhargava | 🔗 aman.bh  | @thedivtagguy</span>"
    )
  ) %>%
  gt_duplicate_column(log_odds_weighted, dupe_name = "point_plot") %>%
  tab_options(row_group.as_column = TRUE) %>%
  fmt_number(columns = c(perc), decimals = 1) %>%
  gt_plt_log(
    column = c(point_plot),
    width = 40,
    palette = c('#4361EE', '#F0F7EE', '#C5283D')
  ) %>%
  cols_label(category = "Type of product",
             point_plot = "Odds of a sale") %>%
  cols_hide(columns = c(
    inner_square_size,
    n,
    product_category,
    emoji,
    perc,
    log_odds_weighted
  )) %>%
  gt_theme_538() 
 
 # unfortunately because of some bug in how gt_group works with row_group.as_column, I was unable to save it in one go as a two column table. From here, I just saved the html file and used puppeteer to take a screenshot.

```

## Creating the dot plot

gtExtras has a nice `gt_plt_point` chart function which inserts SVG charts into table columns. In the first pass, [I tried making my own function to do this](https://gt.albert-rapp.de/fancy_stuff#add-arbitrary-plots-to-your-table) but I couldn't get it to work in the way I wanted on time. So instead, I looked through the source code for the `gt_plt_point` function and adapted it (with very small changes, this is almost the function in its entirety) to work with a fixed scale, my required intercepts and axis labels.

```{r}

add_log_plot <- function(data,
                         palette,
                         add_label,
                         width,
                         vals_range,
                         accuracy) {
  if (data %in% c("NA", "NULL", NA, NULL)) {
    return("<div></div>")
  } else {
    df_in <- dplyr::tibble(x = data,
                           y = 1,
                           color = palette)
    
    val_breaks <-
      seq(from = vals_range[1],
          to = vals_range[2],
          length.out = 7)
    break_labs <-
      ifelse(2 ^ abs(val_breaks) == 1, "Same", paste0(2 ^ abs(val_breaks), "x"))
    
    out_pt_plt <- ggplot(df_in) +
      geom_vline(xintercept = val_breaks,
             color = "#d9d9d9",
             linewidth = 0.25) +
  geom_vline(xintercept = 0,
             color = "#000000",
             linewidth = 0.45) +
      geom_hline(
        yintercept = 1,
        color = "lightgrey",
        linewidth = 0.25,
        linetype = "dotted"
      ) +
      geom_point(
        aes(
          x = .data$x,
          y = .data$y,
          fill = I(.data$color)
        ),
        color = "black",
        size = 3,
        stroke = 0.3,
        shape = 21
      ) +
      theme_void() +
      theme(plot.margin = margin(0, 4, 0, 8, unit = "pt")) +
      coord_cartesian(xlim = vals_range,
                      ylim = c(0.6, 1.2),
                      clip = "off")
    
    if (isTRUE(add_label)) {
      middle_index <-
        ceiling(length(val_breaks) / 2)  # Calculate middle index
      last_label <- break_labs[1]
       if (last_label != "Same") {
    last_label <- paste("1/", last_label, sep = "")
  }

      
      out_pt_plt <- out_pt_plt +
        geom_text(
          aes(x = val_breaks[1], y = .61, label = last_label),
          hjust = 0.5,
          vjust = 0,
          size = 2.5,
          family = "mono",
          color = "black"
        ) +
        geom_text(
          aes(x = val_breaks[length(val_breaks)], y = .61, label = break_labs[length(break_labs)]),
          hjust = 0.5,
          vjust = 0,
          size = 2.5,
          family = "mono",
          color = "black"
        )
    } else {
      out_pt_plt <- out_pt_plt
    }

    out_name <- file.path(tempfile(
      pattern = "file",
      tmpdir = tempdir(),
      fileext = ".svg"
    ))

    ggsave(
      out_name,
      out_pt_plt,
      height = 5,
      width = width,
      dpi = 25.4,
      units = "mm",
      device = "svg"
    )

    img_plot <- readLines(out_name) %>%
      paste0(collapse = "") %>%
      gt::html()

    on.exit(file.remove(out_name), add = TRUE)

    img_plot
  }
}


gt_plt_log <- function(gt_object,
                       column,
                       palette = c("#0091AD", "#EAEBED", "#AF3E4D"),
                       width = 25,
                       scale = 1,
                       range = c(-9, 9),
                       accuracy = 1) {
  
  val_range <- range
  
  gt_object %>%
    text_transform(
      locations = cells_body({
        {
          column
        }
      }),
      fn = function(x) {
        x <- as.double(x) * scale
        n_vals <- 1:length(x)
        
        col_pal <- scales::col_quantile(
          palette = palette,
          domain = val_range,
          reverse = TRUE,
          alpha = TRUE,
          n = 5
        )(x)
        
        add_label <- n_vals %in% c(min(n_vals), max(n_vals))
        
        mapply(add_log_plot,
               x,
               col_pal,
               add_label,
               width,
               list(val_range),
               accuracy,
               SIMPLIFY = FALSE)
      }
    )
}

```

# Abandoned Graveyard

## Creating proportional squares

```{r}

create_square <- function(data, category) {
  data %>%
  filter(product_category == category) %>%
  ggplot() +
  geom_rect(aes(
    xmin = 0,
    xmax = 1,
    ymin = 0,
    ymax = 1
  ), fill = "#CED0CE") +
  geom_rect(aes(
    xmin = 1 - inner_square_size,
    xmax = 1,
    ymin = 0,
    ymax = inner_square_size
  ),
  fill = "#582C4D") +
  coord_fixed() +
  theme_void() +
  theme(
    plot.margin = margin(0, unit = 'pt'),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) 
}

```
